\LARGE \bf Matrix Multiplication-Driven Repulsive Fields for Manipulator Kinematic Obstacle Avoidance

### 2024-03-21 Articles/IROS/root.tex
Mentor Corrections: Round 2

The article presents a new method for obstacle avoidance for robotic manipulators. It uses occupancy voxel grids to calculate repulsive velocities based on the proximity of obstacles. This approach allows for safe and agile navigation in changing environments. Repulsive velocities generated through matrix multiplication are processed using tri-linear interpolation, resulting in smooth and efficient movement away from potential hazards. The integration of this method with inverse kinematic control enables manipulator motion planning and obstacle avoidance capabilities. Simulation results prove the effectiveness of this method in various scenarios, showcasing its potential for local motion planning and obstacle avoidance.

@@

Often used approach is a combination of global and local planning strategies, leveraging global planners for static environment navigation and local planners for adjusting to dynamic changes. This implementation enhances the manipulator's ability to navigate complex environments by combining the comprehensive pathfinding capabilities of global methods with the adaptability and efficiency of local techniques.

@@

Kernel multiplication is employed to ascertain the repulsive velocity components within Cartesian space. We select a segment from the occupation grid, that is centered in the point of interest (POI) on the robot, where we want to calculate the repulsive velocities. The segment needs to be of the same size as the directional matrix, for the cordinate axis in which we are calulating the repulsive velocity. This results in two identically sized 3D matrices: one serving as a "window" into the obstacle grid (\(G_d\)) and the other as the directional kernel (\(W_d\)).

Applying the Hadamard (element-wise) product to the segment of the occupation grid \(G_d\) and the kernel \(W_d\) results in a new matrix. By summing over all values in this resultant matrix, we calculate the repulsive velocity component for that specific direction. This process is independently conducted for each Cartesian coordinate, ultimately yielding the comprehensive repulsive velocity vector \(\vec{v}_i = [ v_{x_i}, v_{y_i}, v_{z_i}]\).

\begin{equation}
	v_d = \sum_{i,j,k} (G \odot W)_{ijk} = \sum_{i}\sum_{j}\sum_{k} g_{\Delta i \Delta j \Delta k} ~  w_{\Delta i \Delta j \Delta k}
	\label{eq:matrix-product}
\end{equation}

@@

The fundamental concept of directional kernels lies in computing the repulsive field individually for each direction within the Cartesian coordinate system.

@@

We introduce two distributions for the primary axis weights, namely gaussian and linear. The selection of weight distribution is contingent on the desired profile of repulsive velocities within the Artificial Potential Field (APF).

@@

\begin{equation}
	w_{\Delta j} =
	\begin{cases} 
		sin(\left| \Delta j \right| \pi / (2 \Delta j_{max})) & \text{if } \Delta j < 0 \text{~or }  \Delta j > 0 \\
		1 & \text{if } \Delta j = 0 \\
	\end{cases}
\end{equation}

@@

\begin{equation}
	w_{\Delta i \Delta j \Delta k} = w_{\Delta i} * w_{\Delta j} * w_{\Delta k}
\end{equation}

@@

The indexes of the eight surrounding cells are calculated by scaling the POI $x,y,z$ by grid resolution and than rounding the position to the nearest lower and upper integer positions (eq.~\ref{eq:floor and ceil}).

\begin{equation}
	\label{eq:floor and ceil}
	\vec{P} =
	\begin{bmatrix}
		X \\
		Y \\
		Z \\
	\end{bmatrix}
	=
	\begin{bmatrix}
		\; \lfloor \vec{p}_{\mathrm{POI}}(1) \times\Delta \mathrm{R} \rfloor \; \lceil \vec{p}_{\mathrm{POI}}(1) \times\Delta \mathrm{R} \rceil \;  \\
		\; \lfloor \vec{p}_{\mathrm{POI}}(2) \times\Delta \mathrm{R} \rfloor \; \lceil \vec{p}_{\mathrm{POI}}(2) \times\Delta \mathrm{R} \rceil \; \\
		\; \lfloor \vec{p}_{\mathrm{POI}}(3) \times\Delta \mathrm{R} \rfloor \; \lceil \vec{p}_{\mathrm{POI}}(3) \times\Delta \mathrm{R} \rceil \; 
	\end{bmatrix}
\end{equation}

Once we got the indexes of the eight surrounding cells of our POI, we calculate the repulsive velocity vectors in each point using kernel matrix multiplication method (seperately for each cartesian direction) (eq.~\ref{eq: calc rep vel}).

\begin{equation}
	\label{eq: calc rep vel}
	\vec{V}rep_{xyz,ijk} = \mathrm{rep}(X[i], Y[j], Z[k]) \quad \forall i, j, k \in \{1, 2\}
\end{equation}

Trilinear interpolation method works on a 3-dimensional regular grid. Before we can start with the interpolation we need to calculate the distance between POI and smaller coordinates of the cells where we calculated the repulsive velocities (eq.~\ref{eq: deltas interp}). The calculated repulsive velocity values are located at the centers of the cells. Therefore, before interpolation, we shift the values of the cells coordinates by half the resolution of the obstacle grid in each direction. That is, the cell coordinates are same as $x,y,z$, cell index and plus a half of grid resolution $\Delta \mathrm{R}$ step.

\begin{equation}
	\label{eq: deltas interp}
	\begin{bmatrix}
		\Delta x \\
		\Delta y \\
		\Delta z		
	\end{bmatrix}
	=
	\begin{bmatrix}
		\frac{\left( P_x - \left( X(1) + \frac{1}{2} \Delta \mathrm{R} \right)  \right)}{\left( X(2) - X(1) \right)} \\
		\frac{\left( P_y - \left( Y(1) + \frac{1}{2} \Delta \mathrm{R} \right)  \right)}{\left( Y(2) - Y(1) \right)} \\
		\frac{\left( P_z - \left( Z(1) + \frac{1}{2} \Delta \mathrm{R} \right)  \right)}{\left( Z(2) - Z(1) \right)} \\
	\end{bmatrix}
\end{equation}

@@

\begin{equation}
\begin{aligned}
	\vec{V}rep_{xyz,jk} = &\vec{V}rep_{xyz,0jk}(1 - \Delta x) \\
	&+ \vec{V}rep_{xyz,1jk} \, \Delta x \quad \forall \, j, k \in \{1, 2\}
\end{aligned}
\end{equation}

@@

\begin{equation}
	\begin{aligned}
		\vec{V}rep_{xyz,k} = &\vec{V}rep_{xyz,0k}(1 - \Delta y) \\
		&+ \vec{V}rep_{xyz,1k} \, \Delta y \quad \forall \, k \in \{1, 2\}
	\end{aligned}
\end{equation}

\begin{equation}
	\label{eq: interp z}
	\vec{V}rep_{xyz} = \vec{V}rep_{xyz,0}(1 - \Delta z) + \vec{V}rep_{xyz,1} \, \Delta z 
\end{equation}

@@

\begin{figure}
	\centering
	\includegraphics[width=1\linewidth]{non_and_interp_4_plots.eps} 
	\caption{Repulsive field generation via convolutional kernels. Left: Original repulsive fields in X (top) and Y (bottom) directions. Right: Corresponding fields using interpolation, showcasing enhanced smoothness.}
	\label{fig:interp-experiment}
\end{figure}

@@

\begin{equation}
	\dot{\vec{q}} = \mathbf{J}^+ \xi_{p} \mathrm{\vec{v}_{att}} + \dot{\vec{q}}_{rep}
\end{equation}

The damped Moore-Penrose pseudo-inverse, \(\mathbf{J}^+~=~\mathbf{J}'(\mathbf{J}\mathbf{J}' + \sigma_{ee} \mathbf{I})^{-1}\), is utilized to mitigate singularity issues and improve numerical stability in inverse kinematics computations. $\xi_{p}$ is the primary task execution slowdown constant. Finally $\dot{\vec{q}}_{rep}$ are the weighted sum of avoidance joint velocities, each transformed into the null space of primary velocities.

@@

\begin{equation} 
	\vec{v} = \frac{\vec{x}_{EE} - \vec{x}_g}{||\vec{x}_{EE} - \vec{x}_g||} \times \frac{\arctan(k_{sigm} \; ||\vec{x}_{EE} - \vec{x}_g||) }{\pi/2}
	\label{eq: v_att}
\end{equation}

In Eq.~\ref{eq: v_att}, $\vec{v}$ specifies the end-effector's (EE) translational velocity directed towards the target, integrating both its direction and magnitude. Here, $\vec{x}_{EE}$ represents the EE's current position, and $\vec{x}_g$ is the target position, both in Cartesian coordinates. The calculation $\frac{\vec{x}_{EE}-\vec{x}_g}{||\vec{x}_{EE}-\vec{x}_g||}$ generates a unit vector pointing towards the target, ensuring targeted movement. The velocity's modulation by the arctangent sigmoid function curtails overshooting by moderating speed in proximity to the target and saturating velocity at a predefined upper limit when distant. The parameter \(k_{sigm}\) in the arctangent function adjusts the curve's steepness, affecting how quickly the end-effector decelerates near the target. A higher \(k_{sigm}\) maintains speed until closer to the target for a sharp deceleration, while a lower value starts slowing down earlier for a gradual approach.

@@

To get the full velocity of the end effector (EE), \( \mathrm{\vec{v}_{att}} \), translational and rotational velocities are scaled using proportional gains and combined \( k_v \) and \( k_\omega \).

\begin{equation}
	\mathrm{\vec{v}_{att}} = 
	\begin{bmatrix}
		k_{v} * \vec{v}   \\
		k_{\omega} * \vec{\omega}
	\end{bmatrix}
	\label{eq:ee_velocity}
\end{equation}

@@

The deceleration factor (eq.~\ref{eq:slowdown}) is determined by the constant \( \kappa_{\text{sec}} \) and the magnitude of the maximum repulsive velocity vectors, which is contingent upon the robot's minimal distance from an obstacle \( v_{\text{Rmax}} \). According to the equation, a small \( v_{\text{Rmax}} \) minimizes the deceleration effect, permitting the uninterrupted execution of the primary task. Conversely, large \( v_{\text{Rmax}} \) boosts the deceleration by reducing the $\xi_{p}$ factor, thereby affording the secondary task increased opportunity for obstacles avoidance motion.

@@

\begin{figure}
	\centering
	\includegraphics[width=1.0\linewidth]{manipulator-kernels-visualization-small.eps} % Replace 'example-image' with the filename of your image
	\caption{Visualization of 3 repulsive Y-direction kernels in POI on the manipulator.}
	\label{fig:kernels}
\end{figure}

@@

During kinematic optimization, repulsive velocities for each point are computed following the methodology described in the previous section (sect. \ref{section:repulsive_vel}). This is achieved by mapping each point into the task space, where the calculation of repulsive velocity is performed. The cartesian mapping for each point comprises a kinematic transformation from the robot's base to the beginning of the segment where the point is located, followed by an additional mapping to the specific point on the segment (eq. \ref{eq:transformations}).  For each point, three components of cartesian velocity are obtained.

@@

Upon obtaining the repulsive velocity for each point, the second norm of each velocity is calculated. For a number of $K$ points where the avoidance velocities are greatest, indicative of closest proximity to an obstacle, the resulting joint velocities that facilitate obstacle avoidance are computed using inverse kinematic equations for the selected points on the manipulator and transformed into the null-space velocities of the primary task (eq.~\ref{eq:point_nullspace}).

@@

The translational velocity Jacobian is calculated for each of the $K$ selected points on the robot. Velocity space can be restricted to just one dimension, on the velocity component that directs movement away from obstacles. Consequently, the Jacobian, correlating joint space velocities \(q\) with the directional velocity \(d_i\), is represented as \(J_{d_i} = \vec{n}_i^T \mathbf{J}_i\). The normal vector, signifying the desired avoidance velocity, is essentially the unit velocity vector, derived from our matrices as \(\vec{n}_{i} = \frac{\vec{v}_{poi i}}{||\vec{v}_{poi i}||}\). This approach streamlines computational processes by avoiding the necessity for complex matrix inversions, transforming the Jacobian from a matrix to a vector of dimensions \(1 \times n\), where \(n\) signifies the count of joints preceding the POI, thus facilitating more straightforward calculations of repulsive velocities.

\begin{equation}
	\dot{q}_{rep} = k_r \sum_{i=1}^{K} \alpha_i J_{d_i}^{+} \left(v_i - J_{d_i} J^{+} \xi_{p} \vec{v}_{att}\right)
\end{equation}

@@

\begin{figure}
	\centering
	\includegraphics[width=0.8\linewidth]{demo-field1c.eps} % Replace 'example-image' with the filename of your image
	\caption{Visualization of the potential field calculated for the entire map using the proposed approach. Top: Obstacle distribution via a 2D heatmap. Center: Induced repulsive fields along X (left) and Y (right) axes. Bottom: Composite vector field showcasing the resultant repulsive velocities for obstacle avoidance.}
	\label{fig:example-field}
\end{figure}

@@

The operation of the potential field is demonstrated on two different manipulator cases. The resolution of our voxel grid for obstacles is \( R=10cm \). The use of interpolation allows us to have a relatively coarse voxel grid, which reduces the memory demand of the space grid and accelerates the computation. We selected \( K = 7 \) POI to observe the distance from obstacles and are uniformly distributed them along the segments and joints, starting from the second joint of the robot to the tip of the manipulator. This distribution begins from the second joint because it is from this point onwards that the manipulator has the capability to avoid obstacles. POIs are indicated by dots on the manipulator in the figures. Near obstacles, vectors emanate from these points, depicting the calculated repulsive velocities at the locations. Throughout the simulation, Euler integration of the calculated joint velocities is performed with a step size of \( T_{step} = 0.1 \) s.

@@

In the second scenario (Fig. \ref{fig:keyframes-3d-ball}, \ref{fig:plots-ball}), our robot operates within a dynamic environment. The proposed method for calculating repulsive velocities proves to be a suitable choice when a local optimization approach is needed to accommodate dynamic changes. In this scenario, a ball moves as governed by the equation \( y_{\text{ball}} = 0.4 + (1.4 - 0.4) \frac{N_{\text{step}}}{75} \),  with \( x_{\text{ball}} = 1.25 \) and \( z_{\text{ball}} = 0.7 \). The primary task is now to maintain the robot's end-effector (EE) at a fixed point, disregarding EE orientation to enhance obstacle avoidance capability. This increases maneuverability, thereby eliminating the need for primary speed task slowdown (\( \xi_{p} = 0 \)). Constants selected for the primary task are \( k_v = 2 \) and \( k_w = 0 \), and for the avoidance task \( k_r = 3 \). Weighting constants for the individual POIs are \( \alpha = \frac{[ \frac{1}{5}, \frac{1}{5}, \frac{1}{5}, \frac{1}{5}, \frac{1}{5}, \frac{1}{5} ]}{10} \).

@@

>>>>

The article presents a new method for obstacle avoidance for robotic manipulators. It uses occupancy voxel grids and convolutional matrix kernels to calculate repulsive velocities based on the proximity of obstacles. This approach allows for safe and agile navigation in changing environments. Repulsive velocities generated through matrix multiplication are processed using tri-linear interpolation, resulting in smooth and efficient movement away from potential hazards. The integration of this method with inverse kinematic control enables manipulator motion planning and obstacle avoidance capabilities. Simulation results prove the effectiveness of this method in various scenarios, showcasing its potential for local motion planning and obstacle avoidance.

@@

Often used approach is a combination~\cite{a15} of global and local planning strategies, leveraging global planners for static environment navigation and local planners for adjusting to dynamic changes. This implementation enhances the manipulator's ability to navigate complex environments by combining the comprehensive pathfinding capabilities of global methods with the adaptability and efficiency of local techniques.

@@

Kernel multiplication is employed to ascertain the repulsive velocity components within Cartesian space. We select a segment from the occupation grid, that is centered in the point of interest (POI) on the robot, where we want to calculate the repulsive velocities. The segment needs to be of the same size as the directional matrix, for the cordinate axis in which we are calulating the repulsive velocity. This results in two identically sized 3D matrices: one serving as a "window" into the obstacle grid $\mathbf{(G_d)}$ and the other as the directional kernel $\mathbf{(W_d)}$.

Applying the Hadamard (element-wise) product to the segment of the occupation grid $\mathbf{(G_d)}$ and the kernel $\mathbf{(W_d)}$ results in a new matrix. By summing over all values in this resultant matrix, we calculate the repulsive velocity component for that specific direction. This process is independently conducted for each Cartesian coordinate, ultimately yielding the comprehensive repulsive velocity vector \(\mathbf{v}_i = [ v_{x_i}, v_{y_i}, v_{z_i}]\).

\begin{equation}
	v_d = \sum_{i,j,k} (\mathbf{G} \odot \mathbf{W})_{ijk} = \sum_{i}\sum_{j}\sum_{k} g_{\Delta i \Delta j \Delta k} ~  w_{\Delta i \Delta j \Delta k}
	\label{eq:matrix-product}
\end{equation}

Where $g_{\Delta i \Delta j \Delta k}$ and $w_{\Delta i \Delta j \Delta k}$ denote $i$-th, $j$-th, $k$-th indexed elements of each matrix.

@@

The fundamental concept of convolutional directional kernels lies in computing the repulsive field individually for each direction within the Cartesian coordinate system.

@@

We introduce two distributions for the primary axis weights, namely Gaussian and linear. The selection of weight distribution is contingent on the desired profile of repulsive velocities within the Artificial Potential Field (APF).

@@

\begin{equation}
	w_{\Delta j} =
	\begin{cases} 
		\sin(\left| \Delta j \right| \pi / (2 \Delta j_{max})) & \text{if } \Delta j < 0 \text{~or }  \Delta j > 0 \\
		1 & \text{if } \Delta j = 0 \\
	\end{cases}
\end{equation}

@@

\begin{equation}
	w_{\Delta i \Delta j \Delta k} = w_{\Delta i} \cdot w_{\Delta j} \cdot w_{\Delta k}
\end{equation}

@@

The indexes of the eight surrounding cells $s=1...8$ are calculated by scaling the selected POI coordinates $x,y,z$ by grid resolution and than rounding the position to the nearest lower and upper integer positions (eq.~(\ref{eq:floor and ceil})).

\begin{equation}
	\label{eq:floor and ceil}
	\vec{P} =
	\begin{bmatrix}
		\mathbf{X} \\
		\mathbf{Y} \\
		\mathbf{Z} \\
	\end{bmatrix}
	=
	\begin{bmatrix}
		\; \lfloor \mathbf{p}_{\mathrm{POI}}(1) \times\Delta \mathrm{R} \rfloor \; \lceil \mathbf{p}_{\mathrm{POI}}(1) \times\Delta \mathrm{R} \rceil \;  \\
		\; \lfloor \mathbf{p}_{\mathrm{POI}}(2) \times\Delta \mathrm{R} \rfloor \; \lceil \mathbf{p}_{\mathrm{POI}}(2) \times\Delta \mathrm{R} \rceil \; \\
		\; \lfloor \mathbf{p}_{\mathrm{POI}}(3) \times\Delta \mathrm{R} \rfloor \; \lceil \mathbf{p}_{\mathrm{POI}}(3) \times\Delta \mathrm{R} \rceil \; 
	\end{bmatrix}
\end{equation}

Once we got the indexes of the eight surrounding cells of our POI, we calculate the repulsive velocity vectors in each point using kernel matrix multiplication method (seperately for each cartesian direction) (eq.~(\ref{eq: calc rep vel})).

\begin{equation}
	\label{eq: calc rep vel}
	\mathbf{V}rep_{xyz,nml} = \mathrm{rep}(\mathbf{X}[n], \mathbf{Y}[m], \mathbf{Z}[l]) \quad \forall n, m, l \in \{1, 2\}
\end{equation}

Trilinear interpolation method works on a 3-dimensional regular grid. Before we can start with the interpolation we need to calculate the distance between POI and smaller coordinates of the cells where we calculated the repulsive velocities (eq.~(\ref{eq: deltas interp})). The calculated repulsive velocity values are located at the centers of the cells. Therefore, before interpolation, we shift the values of the cells coordinates by half the resolution of the obstacle grid in each direction. That is, the cell coordinates are same as $x,y,z$, cell index and plus a half of grid resolution $\Delta \mathrm{R}$ step.

\begin{equation}
	\label{eq: deltas interp}
	\begin{bmatrix}
		\Delta x \\
		\Delta y \\
		\Delta z		
	\end{bmatrix}
	=
	\begin{bmatrix}
		\frac{\left( P_x - \left( \mathbf{X}(1) + \frac{1}{2} \Delta \mathit{R} \right)  \right)}{\left( \mathbf{X}(2) - \mathbf{X}(1) \right)} \\
		\frac{\left( P_y - \left( \mathbf{Y}(1) + \frac{1}{2} \Delta \mathit{R} \right)  \right)}{\left( \mathbf{Y}(2) - \mathbf{Y}(1) \right)} \\
		\frac{\left( P_z - \left( \mathbf{Z}(1) + \frac{1}{2} \Delta \mathit{R} \right)  \right)}{\left( \mathbf{Z}(2) - \mathbf{Z}(1) \right)} \\
	\end{bmatrix}
\end{equation}

@@

\begin{equation}
\begin{aligned}
	\mathbf{V}rep_{xyz,jk} = &\mathbf{V}rep_{xyz,0jk}(1 - \Delta x) \\
	&+ \mathbf{V}rep_{xyz,1jk} \, \Delta x \quad \forall \, j, k \in \{1, 2\}
\end{aligned}
\end{equation}

@@

\begin{equation}
	\begin{aligned}
		\mathbf{V}rep_{xyz,k} = &\mathbf{V}rep_{xyz,0k}(1 - \Delta y) \\
		&+ \mathbf{V}rep_{xyz,1k} \, \Delta y \quad \forall \, k \in \{1, 2\}
	\end{aligned}
\end{equation}

\begin{equation}
	\label{eq: interp z}
	\mathbf{V}rep_{xyz} = \mathbf{V}rep_{xyz,0}(1 - \Delta z) + \mathbf{V}rep_{xyz,1} \, \Delta z 
\end{equation}

@@

\begin{figure}
	\centering
	\includegraphics[width=1\linewidth]{non_and_interp_4_plots.eps} 
	\caption{Repulsive field generation via convolutional kernels. Left: Original repulsive fields in $\mathit{X}$ (top) and $\mathit{Y}$ (bottom) directions. Right: Corresponding fields using interpolation, showcasing enhanced smoothness.}
	\label{fig:interp-experiment}
\end{figure}

@@

\begin{equation}
	\dot{\mathbf{q}} = \mathbf{J}^+ \xi_{p} \mathrm{\mathbf{v}_{att}} + \dot{\mathbf{q}}_{rep}
\end{equation}

The damped Moore-Penrose pseudo-inverse, \(\mathbf{J}^+~=~\mathbf{J}'(\mathbf{J}\mathbf{J}' + \sigma_{ee} \mathbf{I})^{-1}\), is utilized to mitigate singularity issues and improve numerical stability in inverse kinematics computations. $\xi_{p}$ is the primary task execution slowdown constant. Finally $\dot{\mathbf{q}}_{rep}$ are the weighted sum of avoidance joint velocities, each transformed into the null space of primary velocities.

@@

\begin{equation} 
	\mathbf{v} = \frac{\mathbf{x}_{EE} - \mathbf{x}_g}{||\mathbf{x}_{EE} - \mathbf{x}_g||} \times \frac{\arctan(k_{sigm} \; ||\mathbf{x}_{EE} - \mathbf{x}_g||) }{\pi/2}
	\label{eq: v_att}
\end{equation}

In Eq.~(\ref{eq: v_att}), $\mathbf{v}$ specifies the end-effector's (EE) translational velocity directed towards the target, integrating both its direction and magnitude. Here, $\mathbf{x}_{EE}$ represents the EE's current position, and $\mathbf{x}_g$ is the target position, both in Cartesian coordinates. The calculation $\frac{\mathbf{x}_{EE}-\mathbf{x}_g}{||\mathbf{x}_{EE}-\mathbf{x}_g||}$ generates a unit vector pointing towards the target, ensuring targeted movement. The velocity's modulation by the arctangent sigmoid function curtails overshooting by moderating speed in proximity to the target and saturating velocity at a predefined upper limit when distant. The parameter \(k_{sigm}\) in the arctangent function adjusts the curve's steepness, affecting how quickly the end-effector decelerates near the target. A higher \(k_{sigm}\) maintains speed until closer to the target for a sharp deceleration, while a lower value starts slowing down earlier for a gradual approach.

@@

To get the full velocity of the end effector (EE), \( \mathrm{\mathbf{v}_{att}} \), translational and rotational components are multiplied using proportional gains and combined.

\begin{equation}
	\mathrm{\mathbf{v}_{att}} = 
	\begin{bmatrix}
		k_{v} \cdot \mathbf{v}   \\
		k_{\omega} \cdot \mathbf{\omega}
	\end{bmatrix}
	\label{eq:ee_velocity}
\end{equation}

@@

The deceleration factor (eq.~(\ref{eq:slowdown})) is determined by the constant \( \kappa_{\text{sec}} \) and the magnitude of the maximum repulsive velocity vectors, which is contingent upon the robot's minimal distance from an obstacle \( v_{\text{Rmax}} \). According to the equation, a small \( v_{\text{Rmax}} \) minimizes the deceleration effect, permitting the uninterrupted execution of the primary task. Conversely, large \( v_{\text{Rmax}} \) boosts the deceleration by reducing the $\xi_{p}$ factor, thereby affording the secondary task increased opportunity for obstacles avoidance motion.

@@

\begin{figure}
	\centering
	\includegraphics[width=1.0\linewidth]{manipulator-kernels-visualization-small.eps} % Replace 'example-image' with the filename of your image
	\caption{Visualization of only $Y$ repulsive direction kernels ($X$,$Z$ not shown), for 3 seperate POI on the manipulator. The colours of the kernels display their absolute values. We can see, that the magnitude of the kernel values falls when approaching kernel edge.}
	\label{fig:kernels}
\end{figure}

@@

During kinematic optimization, repulsive velocities for each point are computed following the methodology described in the previous section (sect. \ref{section:repulsive_vel}). This is achieved by mapping each point into the task space, where the calculation of repulsive velocity is performed. The cartesian mapping for each point comprises a kinematic transformation from the robot's base to the beginning of the segment where the point is located, followed by an additional mapping to the specific point on the segment (eq.~(\ref{eq:transformations})).  For each point, three components of cartesian velocity are obtained.

@@

Upon obtaining the repulsive velocity for each point, the second norm of each velocity is calculated. For a number of $K$ points where the avoidance velocities are greatest, indicative of closest proximity to an obstacle, the resulting joint velocities that facilitate obstacle avoidance are computed using inverse kinematic equations for the selected points on the manipulator and transformed into the null-space velocities of the primary task (eq.~(\ref{eq:point_nullspace})).

@@

The translational velocity Jacobian is calculated for each of the $K$ selected points on the robot. Velocity space can be restricted to just one dimension, on the velocity component that directs movement away from obstacles. Consequently, the Jacobian, correlating joint space velocities \(q\) with the directional velocity \(d_i\), is represented as \(\mathbf{J}_{d_i} = \mathbf{n}_i^T \mathbf{J}_i\). The normal vector, signifying the desired avoidance velocity, is essentially the unit velocity vector, derived from our matrices as \(\mathbf{n}_{i} = \frac{\mathbf{v}_{poi i}}{||\mathbf{v}_{poi i}||}\). This approach streamlines computational processes by avoiding the necessity for complex matrix inversions, transforming the Jacobian from a matrix to a vector of dimensions \(1 \times n\), where \(n\) signifies the count of joints preceding the POI, thus facilitating more straightforward calculations of repulsive velocities.

\begin{equation}
	\dot{\mathbf{q}}_{rep} = k_r \sum_{i=1}^{K} \alpha_i \mathbf{J}_{d_i}^{+} \left(v_i - \mathbf{J}_{d_i} \mathbf{J}^{+} \xi_{p} \mathbf{v}_{att}\right)
\end{equation}

@@

\begin{figure}
	\centering
	\includegraphics[width=0.8\linewidth]{demo-field1c.eps} % Replace 'example-image' with the filename of your image
	\caption{Visualization of the potential field calculated for the entire map using the proposed approach. Top: Obstacle distribution via a 2D heatmap. Center: Induced repulsive fields along $\mathit{X}$ (left) and $\mathit{Y}$ (right) axes. Bottom: Composite vector field showcasing the resultant repulsive velocities for obstacle avoidance.}
	\label{fig:example-field}
\end{figure}

@@

The operation of the potential field is demonstrated on two different manipulator cases. The resolution of our voxel grid for obstacles is \( R=10\) cm. The use of interpolation allows us to have a relatively coarse voxel grid, which reduces the memory demand of the space grid and accelerates the computation. We selected \( K = 7 \) POI to observe the distance from obstacles and are uniformly distributed them along the segments and joints, starting from the second joint of the robot to the tip of the manipulator. This distribution begins from the second joint because it is from this point onwards that the manipulator has the capability to avoid obstacles. POIs are indicated by dots on the manipulator in the figures. Near obstacles, vectors emanate from these points, depicting the calculated repulsive velocities at the locations. Throughout the simulation, Euler integration of the calculated joint velocities is performed with a step size of \( T_{step} = 0.1 \) s.

@@

In the second scenario (Fig. \ref{fig:keyframes-3d-ball}, \ref{fig:plots-ball}), our robot operates within a dynamic environment. The proposed method for calculating repulsive velocities proves to be a suitable choice when a local optimization approach is needed to accommodate dynamic changes. In this scenario, a ball moves as governed by the equation \( y_{{ball}} = 0.4 + (1.4 - 0.4) \frac{N_{{step}}}{75} \),  with \( x_{{ball}} = 1.25 \) and \( z_{{ball}} = 0.7 \). The primary task is now to maintain the robot's end-effector (EE) at a fixed point, disregarding EE orientation to enhance obstacle avoidance capability. This increases maneuverability, thereby eliminating the need for primary speed task slowdown (\( \xi_{p} = 0 \)). Constants selected for the primary task are \( k_v = 2 \) and \( k_w = 0 \), and for the avoidance task \( k_r = 3 \). Weighting constants for the individual POIs are \( \alpha = \frac{[ \frac{1}{5}, \frac{1}{5}, \frac{1}{5}, \frac{1}{5}, \frac{1}{5}, \frac{1}{5} ]}{10} \).

@@

\bibitem{a15} J. Amiryan and M. Jamzad, "Adaptive motion planning with artificial potential fields using a prior path," in 2015 3rd RSI International Conference on Robotics and Mechatronics (ICROM), Tehran, Iran, 2015, pp. 731-736, doi: 10.1109/ICRoM.2015.7367873. Keywords: Planning; Force; Motion segmentation; Vehicles; Path planning; Mobile robots; Artificial Potential Fields; Autonomous vehicle; Evolutionary Algorithms; Random Sampling; Reactive Motion Planning.
